---
title: "Homework 1"
subtitle: "Due Wednesday Sep 2"
author: "Tsering Dolkar"
date: "8/26/2020"
output: pdf_document
---

# Problem 1 

R is an open source, community built, programming platform.  Not only is there a plethora of useful web based resources, there also exists in-R tutorials.  As per the instructions, I did both of the Primers labeled as The Basics on Rstudio.cloud.

# Problem 2

After having the R environment setup, with the help of a basic understanding of R, I created this Markdown, saved the file to the directory containing the *README.md* file.

### Part A

Although I had a brief experience with R before, I have not studied it from the basics properly and therefore, had problems understanding codes during certain instances in the past. With this class, I hope to have a more innate understanding of the language and to have a certain level of comfort using it in the future. 
Desired learning objectives:

1. Better command and understanding of R 
2. Better understanding of statistical simulations in R
3. Hopefully, learn a bit about SAS 

### Part B

(1) *The density of the sum of two normals*:
$$ If  [Re[z] < 0, \frac{e^{-\frac{z^2}{4}}}{2 \pi}, \int_{-\infty}^\infty \frac{E^{-\frac{y^2}{2}-\frac{1}{2}(-y+z)^2}}{2\pi}dy]$$
(2) *The density of the sum of two Cauchys*:
$$\frac{2}{\pi (-2I+z)(2I+z)}$$
(3) *The density of the sum of two t's with 5 degrees of freedom*:
$$\frac{400 \sqrt{5}(8400+120z^2+z^4)}{3\pi(20+z^2)^5}$$

# Problem 3



# Problem 4

Following is a basic scatter plot:
```{r}
plot(ChickWeight$weight,ChickWeight$Time, cex = 0.2, xlab = "Weight", ylab = "Time", 
     main = "ChickWeight")
```

Following is a histogram:
```{r}
hist(ChickWeight$weight, xlab = "Weight", ylab = "Frequency", 
     main = "ChickWeight")
```
