---
title: "Homework 2"
subtitle: "Due Wednesday Sep 16"
author: "Tsering Dolkar"
date: "9/6/2020"
output: pdf_document
header-includes: 
    \usepackage{tikz,pgfplots}
    \usepackage{fancyhdr}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H', fig.align = "center", echo = TRUE, include = FALSE, eval = TRUE, cache = TRUE)
library(data.table)
library(tidyverse)
```

# Problem 1

I finished the Primers titled “Work with Data” and “Tidy Your Data” on Rstudio cloud.

# Problem 2

Rmarkdown of this HW is opened. 

# Problem 3

While I was working on this Homework, I was not able to commit the changes I made through trial and error and regretted not having a backup. I was going back and forth on which tidy version of data might work for me and for some problems, I lost my code when I made the changes and had to start over on it. As was pointed out in the two links to StackOverflow on why we should use version control, such issues could easily be addressed using version control i.e git.  


# Problem 4

## a.

We are looking for sensory data from five operators from Wu and Hamada's book:
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat> 

First of all, we will get the data from the link above:

```{r sensorydata_import, cache=FALSE, include=TRUE, echo=TRUE}
## getting http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat

url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
sensorydata_raw<- fread(url, data.table = FALSE, fill = TRUE, skip = 2, header = FALSE)
saveRDS(sensorydata_raw, "sensorydata_raw.RDS")
#Saves the object in it's native format without the name: When importing, it is easier for us.
sensorydata_raw <- readRDS("sensorydata_raw.RDS")
```

Need to tidy the data, basic issue is we need items, operators and each of their sensory data on the 10 items are columns.

```{r sensorydata_baseR, cache=FALSE, include=TRUE, echo=TRUE}
#if the first condition is T, return the index value
Na <- which(is.na(sensorydata_raw$V6),arr.ind=T)
df <- cbind(rep(1:10, each = 2), sensorydata_raw[Na,])
df$V6 <- NULL
#create a clone data frame of the raw data, to get a better aligned raw data 
# and not change the original raw data.
new_sensorydata_raw <- sensorydata_raw
new_sensorydata_raw[Na,] <- df
colnames(new_sensorydata_raw) <- c("Item","Operator1","Operator2","Operator3","Operator4","Operator5")
as.data.frame(head(new_sensorydata_raw))
Operator <- stack(new_sensorydata_raw[,2:6])
sensorydata_tidy_br <- data.frame(Item=rep(new_sensorydata_raw$Item, 5),
                                  as.character(Operator[,2]), as.numeric(Operator[,1]))
colnames(sensorydata_tidy_br) <- c('Item', "Operator", "sensorydata")



head(sensorydata_tidy_br)
```

We have converted the dataframes to tidy dataframes using the base functions. Here is a summary of the data:

```{r summary_sensorydata_baseR, include=TRUE,echo=FALSE}
knitr::kable(summary(new_sensorydata_raw))
knitr::kable(summary(sensorydata_tidy_br))
```

Then our boxplot of Item to sensorydata is:
```{r boxplot_sensorydata_baseR, include=TRUE,echo=FALSE}
par(mfrow=c(1,1))
boxplot(sensorydata_tidy_br$sensorydata~sensorydata_tidy_br$Item, col=rainbow(10), xlab = "Item", ylab = "Sensory Data",main="Sensory data from five operators")
```

```{r sensorydata_tidyverse, cache=FALSE, include=TRUE, echo=TRUE}
#if the first condition is T, return the index value
Na <- which(is.na(sensorydata_raw$V6),arr.ind=T)
df <- cbind(rep(1:10, each = 2), sensorydata_raw[Na,])
df$V6 <- NULL
#create a clone data frame of the raw data, to get a better aligned raw data and not
# change the original raw data.
new_sensorydata_raw <- sensorydata_raw
new_sensorydata_raw[Na,] <- df
colnames(new_sensorydata_raw) <- c("Item","Operator1","Operator2","Operator3","Operator4","Operator5")
as.data.frame(head(new_sensorydata_raw))

 sensorydata_tidy_tv <- 
  new_sensorydata_raw %>%
   gather(key = "Operators", value = "sensorydata", Operator1, Operator2, Operator3, Operator4, 
          Operator5)
  
  
head(sensorydata_tidy_tv)
```

We have converted the dataframes to tidy dataframes using the tidyverse functions. Here is a summary of the data:

```{r summary_sensorydata_tidyverse, include=TRUE, echo=FALSE}
knitr::kable(summary(new_sensorydata_raw))
knitr::kable(summary(sensorydata_tidy_tv))
```

Then our boxplot of Item to sensorydata is:
```{r boxplot_sensorydata_tidyverse, include=TRUE,echo=FALSE}
# par(mar = c(10, 10, 0.002, 0.002))
boxplot(sensorydata_tidy_tv$sensorydata~sensorydata_tidy_tv$Item, col=rainbow(10), xlab = "Item", ylab = "sensorydata", main="Sensory data from five operators")
```

## b. 

We are looking at the Gold Medal performance for Olympic Men's Long Jump, year is coded as 1900=0 from Wu and Hamada's book: <http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat> 

First, we will get the data from the link above:

```{r longjumpdata_import, cache=FALSE,include=TRUE, echo=TRUE}
## getting http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat

url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat"

LongJumpData_raw<- fread(url, header = FALSE, fill = TRUE, skip = 1)
saveRDS(LongJumpData_raw, "LongJumpData_raw.RDS")
#Saves the object in it's native format without the name: When importing, it is easier for us.

LongJumpData_raw <- readRDS("LongJumpData_raw.RDS")
```

Need to tidy the data, basic issues are Year and Longjump are columns, need to push them into column.


```{r longjumpdata_baseR, cache=FALSE,include=TRUE, echo=TRUE}
Year <- c(LongJumpData_raw$'V1', LongJumpData_raw$'V3', LongJumpData_raw$'V5', LongJumpData_raw$'V7')
LongJumpData <- c(LongJumpData_raw$V2, LongJumpData_raw$V4, LongJumpData_raw$V6, LongJumpData_raw$V8)
Year <- as.numeric(Year)
LongJumpData <- as.numeric(LongJumpData)
Year <- Year + 1900
longjump_tidy_br <- cbind(Year, LongJumpData)
longjump_tidy_br <- longjump_tidy_br[1:22,]
longjump_tidy_br <- as.data.frame(longjump_tidy_br)

head(longjump_tidy_br)
```

We have converted the dataframes to tidy dataframes using the base functions. Here is a summary of the data:

```{r summary_longjumpdata_baseR, include=TRUE, echo=FALSE}
knitr::kable(summary(longjump_tidy_br))
```

Then our plot of year to Gold Medal performance for Olympic Men’s Long Jump is:
```{r plot_longjump_baseR, include=TRUE,echo=FALSE, warning = FALSE}
plot(longjump_tidy_br$Year, longjump_tidy_br$LongJumpData,cex= 0.4, xlab = "Year", ylab = "LongJump Performance", main = "Gold Medal performance for Olympic Mens Long Jump")
```

```{r longjumpdata_tidyverse, cache=FALSE,include=TRUE, echo=TRUE}
lj_year<- 
  LongJumpData_raw %>%
    gather(key = "Vector2", value = "Year", 'V1','V3','V5','V7', convert = TRUE)%>%
    select(Year)%>% 
    slice(1:(n()-2))%>% #cuts off the NAs at the end
    mutate(Year = 1900 + Year)
lj_data<-
  LongJumpData_raw %>%
    gather(key = "Vector1", value = "LongJumpData", 'V2','V4','V6','V8', convert = TRUE) %>%
    select(LongJumpData)%>% 
    slice(1:(n()-2)) #cuts off the NAs at the end
  
longjump_tidy_tv <- cbind(lj_year, lj_data)
head(longjump_tidy_tv)
```
We have converted the dataframes to tidy dataframes using the tidyverse functions. Here is a summary of the data:

```{r summary_longjumpdata_tidyverse, include=TRUE, echo=FALSE}
knitr::kable(summary(longjump_tidy_tv))
```

Then our plot of year to Gold Medal performance for Olympic Men’s Long Jump is:
```{r plot_longjump_tidyverse, include=TRUE,echo=FALSE,warning = FALSE}
plot(longjump_tidy_tv$Year, longjump_tidy_tv$LongJumpData,cex= 0.4, xlab = "Year", ylab = "LongJump Performance", main = "Gold Medal performance for Olympic Mens Long Jump")
```

## c. 

We are looking at brain weight (g) and body weight (kg) for 62 species from Wu and Hamada's book: <http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat> 

First, we will get the data from the link above:

```{r brainandbodyweight_import, cache=FALSE,include=TRUE, echo=TRUE}
## getting http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat

url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat"

brainandbodyweight_raw<- fread(url, header = FALSE, fill = TRUE, skip = 1)
saveRDS(brainandbodyweight_raw, "brainandbodyweight_raw.RDS")
#Saves the object in it's native format without the name: When importing, it is easier for us.

brainandbodyweight_raw <- readRDS("brainandbodyweight_raw.RDS")
```


Need to tidy the data, basic issues are brain weight and body weight are columns, need to push them into column.

```{r brainandbodyweight_baseR, cache=FALSE,include=TRUE, echo=TRUE}
bodyweight <- c(brainandbodyweight_raw$'V1', brainandbodyweight_raw$'V3', brainandbodyweight_raw$'V5')
brainweight <- c(brainandbodyweight_raw$V2, brainandbodyweight_raw$V4, brainandbodyweight_raw$V6)
bodyweight <- as.numeric(bodyweight)
brainweight <- as.numeric(brainweight)
brainandbodyweight_tidy_br <- cbind(bodyweight, brainweight)
brainandbodyweight_tidy_br <- brainandbodyweight_tidy_br[1:62,]
brainandbodyweight_tidy_br <- as.data.frame(brainandbodyweight_tidy_br)

head(brainandbodyweight_tidy_br)

```

We have converted the dataframes to tidy dataframes using the base functions. Here is a summary of the data:

```{r summary_brainandbodyweight_baseR, include=TRUE, echo=FALSE}
knitr::kable(summary(brainandbodyweight_tidy_br))
```

Then our plot of Brain weight (g) and body weight (kg) for 62 species:
```{r plot_brainandbodyweight_baseR, include=TRUE,echo=FALSE}

plot(brainandbodyweight_tidy_br$bodyweight, brainandbodyweight_tidy_br$brainweight, xlab = "Body weight (g)", ylab = "Brain weight (kg)", main = "Brain weight (g) and body weight (kg) for 62 species")
```

```{r brainandbodyweight_tidyverse, cache=FALSE,include=TRUE, echo=TRUE}
bodyweight<- 
  brainandbodyweight_raw %>%
    gather(key = "Vector2", value = "bodyweight", V1, V3, V5, convert = TRUE)%>%
    select(bodyweight)%>%
    slice(1:(n()-1)) #cuts off the NAs at the end

brainweight<-
  brainandbodyweight_raw %>%
    gather(key = "Vector1", value = "brainweight", V2, V4, V6, convert = TRUE)%>%
    select(brainweight)%>%
    slice(1:(n()-1)) #cuts off the NAs at the end
  
brainandbodyweight_tidy_tv <- cbind(bodyweight, brainweight)
  

head(brainandbodyweight_tidy_tv)
```
We have converted the dataframes to tidy dataframes using the tidyverse functions. Here is a summary of the data:

```{r summary_brainandbodyweight_tidyverse, include=TRUE, echo=FALSE}
knitr::kable(summary(brainandbodyweight_tidy_tv))
```

Then our plot of Brain weight (g) and body weight (kg) for 62 species:
```{r plot_brainandbodyweight_tidyverse, include=TRUE,echo=FALSE}

plot(brainandbodyweight_tidy_tv$bodyweight, brainandbodyweight_tidy_tv$brainweight,xlab = "Body weight (g)", ylab = "Brain weight (kg)", main = "Brain weight (g) and body weight (kg) for 62 species")
```

## d. 

We will look to triplicate measurements of tomato yield for two varieties of tomatos at three planting densities from Wu and Hamada's book: <http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat> 

First, we will get the data from the link above:

```{r tomato_import, cache=FALSE,include=TRUE, echo=TRUE}
## getting http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat

url <- "http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat"

tomato_raw<- fread(url)
saveRDS(tomato_raw, "tomato_raw.RDS")
#Saves the object in it's native format without the name: When importing, it is easier for us.

tomato_raw <- readRDS("tomato_raw.RDS")
```

Need to tidy the data, basic issue is we need a column of different densities, two column of measurements for each type of tomato and a column called M number to keep track of which of 3 measurement for each density we are looking at. 

```{r tomato_baseR, cache=FALSE, include=TRUE, echo=TRUE }
#create a dataframe called tomato_tidy where there is 3 of the tomatovariety, the string of
# their measurements, and densities.
tomato_tidy_br <- data.frame(TomatoVariety=rep(tomato_raw$V1, 3),
                          stack(tomato_raw[,-1]))
colnames(tomato_tidy_br) <- c("TomatoVariety", "V1", "Density")

# we will now try to separate the string of measurement into individual measurements
M <- data.frame(strsplit(head(tomato_tidy_br$V1), ','))
colnames(M) <- c("1", "2", "3", "4", "5", "6")
rownames(M) <- c("M1", "M2", "M3")
M <- as.data.frame(t(M))

#Separate the measurements into columns of measurements for Tomato type 1 and type 2
Type1 <- as.data.frame(stack(M[c(1,3,5),]))
colnames(Type1) <- c("Ife1", "M_number")
# Since the measurement number is redundant, remove one column out of two M numbers
Type1$M_number <- NULL
Type2 <- as.data.frame(stack(M[c(2,4,6),]))
colnames(Type2) <- c("PusaEarlyDwarf","M_number")
# bind the two types together into one dataframe
df <- as.data.frame(cbind(Type1, Type2))
#change the type of the variables to reflect true type.
df$Ife1 = as.numeric(df$Ife1)
df$PusaEarlyDwarf=as.numeric(df$PusaEarlyDwarf)
df1 <- as.data.frame(subset(df, select = c(Ife1, PusaEarlyDwarf)))
df2 <- as.data.frame(subset(df, select = M_number))

#create a dataframe with column of density values and dataframe with the rest of the information
subset.df = data.frame(Density=rep(c(1000,2000,3000), 3))
subset.df <- as.data.frame(subset.df)
new.df <- data.frame(cbind(df2, df1))

# bind the density, m number, and measurements of the two types of tomatoes and reoder the
# rows of the dataframe to look more like the tidyverse dataframe(since I did that first)
# for easy comparison reason.
tomato_tidy_br <- cbind(subset.df, new.df)
colnames(tomato_tidy_br) <- c("Density", "M number", 'Ife1', "PusaEarlyDwarf")
tomato_tidy1 <- data.frame(tomato_tidy_br[c(1,4,7),])
tomato_tidy2 <- data.frame(tomato_tidy_br[c(2,5,8),])
tomato_tidy3 <- data.frame(tomato_tidy_br[c(3,6,9),])
tomato_tidy_br <- data.frame(rbind(tomato_tidy1,tomato_tidy2,tomato_tidy3))
#the automated row index was showing up as out of order. For aesthetics, I removed it.
rownames(tomato_tidy_br) <- NULL


head(tomato_tidy_br)
```

We have converted the dataframes to tidy dataframes using the base functions. Here is a summary of the data:

```{r summary_tomato_baseR, include=TRUE, echo=FALSE}
knitr::kable(summary(tomato_tidy_br))
```

Then our plot of density to measurements of tomato yield for two varieties of tomatoes:
```{r plot_tomato_baseR, include=TRUE,echo=FALSE}
par(mfrow=c(1,2))


boxplot(tomato_tidy_br$Ife1~tomato_tidy_br$Density, col=blues9, xlab = "Density", ylab = "Ife1")
boxplot(tomato_tidy_br$PusaEarlyDwarf~tomato_tidy_br$Density, col = rainbow(1), xlab = "Density", ylab = "PusaEarlyDwarf")
title(main=print("Measurements of tomato yield for two types of tomatoes"),outer=T, line=-2)
```

```{r tomato_tidyverse, cache = FALSE,include=TRUE, echo=TRUE}
tomato_tidy_tv <- 
  tomato_raw %>%
  gather(key = "Density", value = "Measurements", -V1, convert = TRUE) %>%
  rename(TomatoVariety = V1) %>%
  separate(Measurements, into = c("M1", "M2", "M3"), sep = ',', convert = TRUE) %>%
  gather(key = "M number", value = "Measurements", 3,4,5)%>% 
  spread(key = TomatoVariety, value = Measurements)

colnames(tomato_tidy_tv) <- c('Density', "M.number", "Ife1", "PusaEarlyDwarf")

head(tomato_tidy_tv)
```
We have converted the dataframes to tidy dataframes using the tidyverse functions.Here is a summary of the data:

```{r summary_tomato_tidyverse, include=TRUE, echo=FALSE}
knitr::kable(summary(tomato_tidy_tv))
```

Then our plot of density to measurements of tomato yield for two varieties of tomatoes:
```{r plot_tomato_tidyverse, include=TRUE,echo=FALSE}
par(mfrow=c(1,2))


boxplot(tomato_tidy_tv$Ife1~tomato_tidy_tv$Density, col=blues9, xlab = "Density", ylab = "Ife1")
boxplot(tomato_tidy_tv$PusaEarlyDwarf~tomato_tidy_tv$Density, col = rainbow(1), xlab = "Density", ylab = "PusaEarlyDwarf")
title(main=print("Measurements of tomato yield for two types of tomatoes"),outer=T, line=-2)
```
## Problem 5

Finish this homework by pushing your changes to your repo.  In general, your workflow for this should be:  

1. git pull -- to make sure you have the most recent repo  
2. In R: do some work  
3. git add -- this tells git to track new files  
4. git commit -- make message INFORMATIVE and USEFUL  
5. git push -- this pushes your local changes to the repo  

If you have difficulty with steps 1-5, git is not correctly or completely setup.  See me for help.
