---
title: "HW3_tdolkar"
author: "Tsering Dolkar"
date: "9/24/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: Due Wednesday Sep 30
header-includes: \usepackage{tikz,pgfplots} \usepackage{fancyhdr}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H', fig.align = "center", echo = TRUE, include = FALSE, eval = TRUE, cache = TRUE)
library(data.table)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(microbenchmark)
library(multipanelfigure)
```

## Problem 1
I did primer on Rstudio cloud.

## Problem 2
Created the Rmd file

## Problem 3
In the lecture, there were two links to programming style guides.
\begin{itemize}
\item Google's R style guide:      https://google.github.io/styleguide/Rguide.xml\\
\item Hadley Wickam’s Style Guide: http://r-pkgs.had.co.nz/style.html
\end{itemize}
While the two styling had some differences, it mainly encouraged coders to write informative comments between lines of code,
to choose informative and original variable/function names, and to make the code neat and readable. I felt that these were advices that will not only be helpful in understanding code while working on projects with other people, but also when I revisit my codes after some time has passed.

## Problem 4
Good programming practices start with this homework.  In the last homework, you imported, munged, cleaned and summarized datasets from Wu and Hamada's _Experiments: Planning, Design and Analysis_.

## Problem 5

```{r data_import}
#A dataset which has multiple repeated measurements from two devices 
#(dev1 and dev2) by thirteen Observers.
observations <- readRDS("~/Downloads/Stat Progr Packages/STAT5014_tdolkar/HW3_data.rds")
```

```{r function_summary, include = T}
summarise_data <- function(our_data){
  #function to calculate summary of a dataframe, returns a vector of the summary
  observer_summary <- double(length = 5)
  for(i in 1:5){
    if(i == 1)
      observer_summary[i] <- mean(our_data[,1])
    if(i == 2)
      observer_summary[i] <- mean(our_data[,2])
    if(i == 3)
      observer_summary[i] <- sd(our_data[,1])
    if(i == 4)
      observer_summary[i] <- sd(our_data[,2])
    if(i == 5)
      observer_summary[i] <- cor(our_data[,1], our_data[,2])
  }
  return(observer_summary)
}
```


```{r mine, include=T}
# We'll find the summary for each observer:
# We will be returned a vector of length 5 for each of 13 observers from the function 
# summarise_data such that 
# 1. mean of dev 1
# 2. mean of dev 2
# 3. standard dev of dev 1
# 4. standard dev of dev 2
# 5. correlation between dev 1 and 2
# are the values respectively in a row.

#separate the data into vectors to make it less confusing to work with first
observer <- observations$Observer
dev1 <- observations$dev1
dev2 <- observations$dev2

#initialize vectors we will need:
dev1_by_observer <- double(0)
dev2_by_observer <- double(0)
summary_statistics <- data.frame()

#we have a nested for loop here. The outer for loop keeps track of observer 1 to 13 and
#the inside for loop looks for all the data by the said observer in the dataset from top to bottom
#once.
for(i in 1:13){
   track_observer <- i
   for(j in 1:length(observer)){
     if(track_observer == observer[j]){
     dev1_by_observer <- c(dev1_by_observer, dev1[j])
     dev2_by_observer <- c(dev2_by_observer, dev2[j])
     }
   }
   raw_data <- data.frame(dev1_by_observer, dev2_by_observer)
   colnames(raw_data) <- c("dev1", "dev2")
   summary_statistics_each_observer <- cbind(rep(i, 5), summarise_data(raw_data))
   summary_statistics <- rbind(summary_statistics, summary_statistics_each_observer)
}
summary_statistics <- data.frame(rep(c("Mean_dev1", "Mean_dev2", "SD_dev1", "SD_dev2", 
                                       "Cor_dev1_dev2"), 13), summary_statistics)
colnames(summary_statistics) <- c('V1','Observer','V3')
summary_statistics <- summary_statistics %>%
                         spread(key = V1, value = V3)
# reorder by column name
summary_statistics <- summary_statistics[c("Observer", "Mean_dev1", "Mean_dev2", 
                                                   "SD_dev1", "SD_dev2", "Cor_dev1_dev2")]
# summary_statistics <- kable(summary_statistics)
print(summary_statistics)
```
From this, it seems that device 2 has a lot bigger standard deviation than device 1 and device 1 and device 2 has negative and weak correlation. 

```{r boxplot_summary, include = TRUE}

observations$Observer <- as.factor(observations$Observer)

dev1 <- observations %>%
  ggplot(aes(x = Observer, y = dev1, colour = Observer)) +
  geom_boxplot() +
  labs(title = "boxplot summary of dev 1", x="Observer", y = "Dev 1") +
  theme_classic()

dev2 <- observations %>%
  ggplot(aes(x = Observer, y = dev2, colour = Observer)) +
  geom_boxplot() +
  labs(title = "boxplot summary of dev 2", x="Observer", y = "Dev 2") +
  theme_classic()

figure1 <- multi_panel_figure(columns = 1, rows = 2, panel_label_type = "none")
figure1 %<>%
  fill_panel(dev1, column = 1, row = 1) %<>%
  fill_panel(dev2, column = 1, row = 2) 
figure1
```
From these plots, we can see that the deviation from median or the IQR is much bigger for device 2 than device 1 for almost all observers. It also seems that the median is close to the mean of the device. Perhaps the distribution for the two devices is normal distribution?

```{r violin_plot_summary, include = TRUE}

dev1 <- observations %>%
  ggplot(aes(x = Observer, y = dev1, colour = Observer)) +
  geom_violin(trim = F) +
  geom_boxplot(width=0.1) +
  stat_summary(fun.y=median, geom="point", size=2, color="red") +
  labs(title = "violinplot summary of dev 1", x="Observer", y = "Dev 1") +
  theme_classic()

dev2 <- observations %>%
  ggplot(aes(x = Observer, y = dev2, colour = Observer)) +
  geom_violin(trim = F) + 
  geom_boxplot(width=0.1) +
  stat_summary(fun.y=median, geom="point", size=2, color="red") +
  labs(title = "violinplot summary of dev 2", x="Observer", y = "Dev 2") +
  theme_classic()

figure1 <- multi_panel_figure(columns = 1, rows = 2, panel_label_type = "none")
figure1 %<>%
  fill_panel(dev1, column = 1, row = 1) %<>%
  fill_panel(dev2, column = 1, row = 2) 
figure1
```
For the most part, it seems to give us the same information as we had in boxplot. The red dot on this plot is our mean. So therefore, we know that our median and mean is the same value. This implies dev1 and dev2 data is normal in distribution. One extra information from this plot is also that the shape of the individual violins displays frequencies of values. 

```{r scatter_plot_summary, include=TRUE,echo=FALSE}
par(mfrow=c(1,1))
observations$Observer <- as.factor(observations$Observer)
observations %>%
  ggplot(aes(x=dev1,y=dev2)) + geom_point() + facet_wrap(Observer~.)

```
Combining the scatter plot with the summary statistics, it is very shocking to see. With similar mean, and standard deviation for the 10 observers, I thought the scatterplot would be more similar if not the same.\\   

As you approach data analysis, what things should you do in the "Exploratory Data Analysis" portion of a project to avoid embarrassment from making erroneos conclusions?\\

We should plot our data in several different ways to glean as much information as we can from all of it together instead of just plotting something once and drawing conclusion from it.

## Problem 6

```{r Riemann_Sum, include=T}
R_sum <- function(a, b, n){
#takes the interval(a,b) and partitions it in n
#pieces. Then returns the sum of all the area.
#the bigger the n value, the closer R_sum value
#gets to integral value.
  area_rectangle <- double()
  dx <- (b - a)/n
  for(i in 1:n){
    x <- a + dx*i
    f_x <- exp((-x^2)/2)
    area_rectangle[i] <- dx * f_x
  }
  return(sum(area_rectangle))
}
```

```{r Riemann_call, include=T}
how_many_parts <- seq(from = 600000, to = 1600000, by = 100000)
a <- 0
b <- 1
RS <- double()
track_RS <- data.frame()
report <- data.frame()

integrand <- function(x) {exp((-x^2)/2)}
integral_value <- integrate(integrand, lower = 0, upper = 1)
integral_value <- integral_value$value

# calls the function and calculates the Riemann of sum for each 
# increasing number of total partition.
for(i in how_many_parts){
  RS <- R_sum(a, b, i) 
  slice_width <- ((b-a)/i)
  track_RS <- rbind(track_RS, c(slice_width, RS))
  # if riemann sum is within e^(-6) points from the integral value, store it in
  # dataframe report.
  if(abs(RS - integral_value) <= 1e-6){
  report <- rbind(report, c(slice_width, RS))
  }
}
colnames(track_RS) <- c("slice_width", "Riemann_Sum")
if(length(report) > 0){
  colnames(report) <- c("slice_width", "Riemann_Sum")
}
print(report)
summary(report)
```

The width necessary to obtain an answer within $1e^-6$ of the analytical solution for 
me was the min value of slice_width. At that slice_width, I get the min value of Riemann_Sum. I could make the answer more accurate by lowering the by value in seq(how_many_parts)[for example, by = 1]. However, that slows down the compile time on my personal computer to hours. 

## Problem 7

```{r Newton_Method, include=T}
N_method <- function(x1 = -5){
  #function takes in a first approximation of the root x0 of the function 3^x-sin(x)+cos(5*x)
  #function uses Newton's method of approximate the root of the function
  #function terminates when successive estimates are within 1e-04 of each other
  seq <- 100
  x <- double(0)
  x[1] <-  x1
  # f <- 3^x- sin(x) + cos(5*x)
  # f_diff <- 3^x * log(3) - cos(x) - sin(5 * x) * 5
  for(i in 2:seq){
    if(3^x[i-1] * log(3) - cos(x[i-1]) - sin(5 * x[i-1]) * 5 == 0){
    stop("choose a new starting place")
    }
    x[i] <- x[i-1] - ((3^x[i-1]- sin(x[i-1]) + cos(5*x[i-1]))/
                          (3^x[i-1] * log(3) - cos(x[i-1]) - sin(5 * x[i-1]) * 5))
    if (abs(x[i]-x[i-1])<1e-04){
      break
    }
  }
  #For the initial estimate of:
  print(x[1])
  #Xn approaches 0 at: 
  print(tail(x,1))
  return(plot(x, type = 'b', col = 'purple'))
}
```


```{r Newton_M_call, include=T}
# we have: 
#f <- expression(3^x- sin(x) + cos(5*x))
init_x <- c(-28, -24, -18, -6)
par(mfrow = c(2,2))
for(i in 1:length(init_x)){
  p = N_method(init_x[i])
}
p
```

## Problem 8

```{r compare, include=T}
#Given simulated data from question 
X <- cbind(rep(1,100),rep.int(1:10,time=10))
beta <- c(4,5)
y <- X%*%beta + rnorm(100)

# precalculation and vectors for SST using for_loop:
y_bar <- mean(y)
SST <- double(0)

# required values and matrices for SST using matrix operation:
n <- 100
J <- matrix(1, nrow = n, ncol = n)
I <- diag(x = 1, nrow = n, ncol = n)

times <- microbenchmark(SST1 <- {for(i in 1:length(y)){ SST[i] <- (y[i] - y_bar)^2};sum(SST)},
               SST2 <- {t(y)%*%(I - (1/n)*J)%*%y}, times = 100, unit = "ms")
print(times)
print(SST1)
print(SST2)
```




